---
title: "Market Clusters"
author: "Faizan Khalid Mohsin"
date: "July 11, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```

# Original Code


```{r library, warning=F}

library(statar)
library(scales)
library(ggplot2)
library(tidyverse)
library(factoextra)
library(ggpubr)
library(knitr)
library(cluster)
library(stats)
```



```{r }

#reading data
Jewl_csr <- read_csv("jewelry_customers.csv")
#exploring head                     
head(Jewl_csr)

#exploring the data and the features prior clustering

any(is.na(Jewl_csr)) # = FALSE meaning no missing data

# scatter plot visulization based on Age and income with color identifying the saving, very clear clusters 
Pl1 <- ggplot(Jewl_csr, aes(Age, Income)) +
  geom_point(aes(color = Savings), alpha= 0.5, size = 3) + 
  scale_color_gradient(low= "red", high = "green")   
Pl1

```


```{r}

#variation between features
var(Jewl_csr$Age)
var(Jewl_csr$Income)
var(Jewl_csr$SpendingScore)  # there is a need for scaling

```



```{r}


#the function appear to scale & unscale the data properly to original state
scaled <- scale(Jewl_csr)
head(scaled, 10)
#Unscaled <- unscale(Jewl_csr)
#head(Unscaled, 10)   # correct data unscaled



```



```{r}
#clusters..............

#1. K-means

# determine the number of clusteres by Elbow plot method
#creating function for wss plot

set.seed(1234)
# Compute and plot wss for k = 2 to k = 15.
k.max <- 15
data <- scaled
wss <- sapply(1:k.max, 
              function(k){kmeans(data, k, nstart=50,iter.max = 15 )$tot.withinss})
wss
plot(1:k.max, wss,
     type="b", pch = 19, frame = FALSE, 
     xlab="Number of clusters K",
     ylab="Total within-clusters sum of squares")  # number 5 appears to be optimal also consistent with Pl1 plot on original data

km <- kmeans(scaled, 5)
#autoplot(km,scaled,frame = TRUE)
km  #examine centers and features of the clusters, centers appears to be diffrent

```



# My Code

## K-means Clustering

### Summary Discriptions of Clusters

```{r }

jewl_data = Jewl_csr
jewl_data$cluster = as.factor(km$cluster)



cluster_summary = jewl_data %>% group_by(cluster) %>% summarise_all(.funs = c(mean="mean", median="median", sd = "sd") )

kable(cluster_summary[, 1:5])



```

```{r}
kable(cluster_summary[, c(1, 6:9)])
```

```{r}
kable(cluster_summary[, c(1, 10:13)])
```



```{r}
cluster = jewl_data$cluster
```



## K-means Clusting Visualizations


```{r}

annotate_figure(
ggarrange(
ggplot(jewl_data, aes(x = Income, y = SpendingScore)) + geom_point(aes(color = cluster)), 
ggplot(jewl_data, aes(x = Age, y = SpendingScore)) + geom_point(aes(color = cluster)), 
nrow  = 1, ncol = 2, common.legend = TRUE, legend = "right"), 
top = text_grob("Plotting K-Means Clusters"))


```


```{r}
annotate_figure(
ggarrange(
ggplot(jewl_data, aes(x = Income, y = Savings)) + geom_point(aes(color = cluster)), 
ggplot(jewl_data, aes(x = Age, y = Savings)) + geom_point(aes(color = cluster)),
ggplot(jewl_data, aes(x = Savings, y = SpendingScore)) + geom_point(aes(color = cluster)), 
ggplot(jewl_data, aes(x = Age, y = Income)) + geom_point(aes(color = cluster)),
nrow  = 2, ncol = 2, common.legend = TRUE, legend = "right"), 
top = text_grob("Plotting K-Means Clusters"))
```



```{r}

library(ggpubr)
ggscatter(jewl_data, x = "Income", y = "SpendingScore",
          color = "cluster", palette = "npg",
          shape = "cluster", ellipse = TRUE, 
          mean.point = TRUE, star.plot = TRUE,
          ggtheme = theme_minimal())
```



```{r}

# Grouped Scatter plot with marginal density plots
# ggscatterhist(
#   jewl_data, x = "Income", y = "SpendingScore",
#   color = "cluster", size = 5, alpha = 0.6,
#   palette = get_palette(c("#00AFBB", "#FC4E07"), 5)), 
#   margin.params = list(fill = "cluster", color = "black", size = 0.2)
# ) 
#  

ggscatterhist(
  jewl_data, x = "Income", y = "SpendingScore",
  color = "cluster", size = 3, alpha = 0.6,
  margin.params = list(fill = "cluster", color = "black", size = 0.2)
  )

```



```{r}
jewl_data$cluster = as.numeric(jewl_data$cluster)

fviz_cluster(list(data = jewl_data, cluster = cluster))
```




#  Hierarchical Clustering



```{r}
scaled_jewl_data = as.data.frame(scale(Jewl_csr))
fviz_nbclust(scaled_jewl_data, FUN = hcut, method = "wss")

```


```{r}
fviz_nbclust(scaled_jewl_data, FUN = hcut, method = "silhouette")
```


```{r}
gap_stat <- clusGap(scaled_jewl_data, FUN = hcut, nstart = 25, K.max = 10, B = 50)
fviz_gap_stat(gap_stat)

```



### Based on wws and gap statistics optimal number of clusters should be 4. However, we will first do Hierarchical Clustering using 5 clusters and then for 4 clusters.

## Hierarchical Clustering using K = 5


```{r}
df = scaled_jewl_data

# methods to assess
m <- c( "average", "single", "complete", "ward")
names(m) <- c( "average", "single", "complete", "ward")

# function to compute coefficient
ac <- function(x) {
  agnes(df, method = x)$ac
}

map_dbl(m, ac)
```



```{r}
# Dissimilarity matrix
d <- dist(df, method = "euclidean")

# Hierarchical clustering using ward method
hc2 <- hclust(d, method = "ward.D2" )
```


```{r}
# Cut tree into 4 groups
sub_grp <- cutree(hc2, k = 5)

# Number of members in each cluster
table(sub_grp)
```




```{r}
jewl_data = jewl_data %>% mutate(h_cluster = sub_grp)
```



```{r}

plot(hc2, cex = 0.6)
rect.hclust(hc2, k = 5, border = 2:5)
```




### Summary Discriptions of Hierarchical Clusters

```{r }

jewl_data$h_cluster = as.factor(jewl_data$h_cluster)


cluster_summary = jewl_data[, c(1:4, 6)] %>% group_by(h_cluster) %>% summarise_all(.funs = c(mean="mean", median="median", sd = "sd") )

kable(cluster_summary[, 1:5])

```

```{r}
kable(cluster_summary[, c(1, 6:9)])
```

```{r}
kable(cluster_summary[, c(1, 10:13)])
```



```{r}
cluster = jewl_data$h_cluster
```



## Hierarchical Clustering Visualizations

```{r}

annotate_figure(
ggarrange(
ggplot(jewl_data, aes(x = Income, y = SpendingScore)) + geom_point(aes(color = h_cluster)), 
ggplot(jewl_data, aes(x = Age, y = SpendingScore)) + geom_point(aes(color = h_cluster)), 
nrow  = 1, ncol = 2, common.legend = TRUE, legend = "right"), 
top = text_grob("Plotting Hierarchical Clusters"))


```


```{r}
annotate_figure(
ggarrange(
ggplot(jewl_data, aes(x = Income, y = Savings)) + geom_point(aes(color = h_cluster)), 
ggplot(jewl_data, aes(x = Age, y = Savings)) + geom_point(aes(color = h_cluster)),
ggplot(jewl_data, aes(x = Savings, y = SpendingScore)) + geom_point(aes(color = h_cluster)), 
ggplot(jewl_data, aes(x = Age, y = Income)) + geom_point(aes(color = h_cluster)),
nrow  = 2, ncol = 2, common.legend = TRUE, legend = "right"), 
top = text_grob("Plotting Hierarchical Clusters"))
```


```{r}

library(ggpubr)
ggscatter(jewl_data, x = "Income", y = "SpendingScore",
          color = "h_cluster", palette = "npg",
          shape = "h_cluster", ellipse = TRUE, 
          mean.point = TRUE, star.plot = TRUE,
          ggtheme = theme_minimal())
```



```{r}

# Grouped Scatter plot with marginal density plots
# ggscatterhist(
#   jewl_data, x = "Income", y = "SpendingScore",
#   color = "cluster", size = 5, alpha = 0.6,
#   palette = get_palette(c("#00AFBB", "#FC4E07"), 5)), 
#   margin.params = list(fill = "cluster", color = "black", size = 0.2)
# ) 
#  

ggscatterhist(
  jewl_data, x = "Income", y = "SpendingScore",
  color = "h_cluster", size = 3, alpha = 0.6,
  margin.params = list(fill = "h_cluster", color = "black", size = 0.2)
  )

```



```{r}

jewl_data$h_cluster = as.numeric(jewl_data$h_cluster)
h_cluster = as.numeric(jewl_data$h_cluster)
fviz_cluster(list(data = jewl_data, cluster = h_cluster))
```



<!-- # Hierarchical Clustering using K = 4 ( 4 Clusters) -->


<!-- ```{r} -->
<!-- # Cut tree into 4 groups -->
<!-- sub_grp <- cutree(hc2, k = 4) -->

<!-- # Number of members in each cluster -->
<!-- table(sub_grp) -->
<!-- ``` -->




<!-- ```{r} -->
<!-- jewl_data = jewl_data %>% mutate(h_cluster = sub_grp) -->
<!-- ``` -->



<!-- ```{r} -->
<!-- plot(hc2, cex = 0.6) -->
<!-- rect.hclust(hc2, k = 4, border = 2:5) -->
<!-- ``` -->




<!-- ### Summary Discriptions of Hierarchical Clusters -->

<!-- ```{r } -->

<!-- jewl_data$h_cluster = as.factor(jewl_data$h_cluster) -->


<!-- cluster_summary = jewl_data[, c(1:4, 6)] %>% group_by(h_cluster) %>% summarise_all(.funs = c(mean="mean", median="median", sd = "sd") ) -->

<!-- kable(cluster_summary[, 1:5]) -->

<!-- ``` -->

<!-- ```{r} -->
<!-- kable(cluster_summary[, c(1, 6:9)]) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- kable(cluster_summary[, c(1, 10:13)]) -->
<!-- ``` -->



<!-- ```{r} -->
<!-- cluster = jewl_data$h_cluster -->
<!-- ``` -->





<!-- ## Hierarchical Clustering Visualizations -->


<!-- ```{r} -->

<!-- annotate_figure( -->
<!-- ggarrange( -->
<!-- ggplot(jewl_data, aes(x = Income, y = SpendingScore)) + geom_point(aes(color = h_cluster)),  -->
<!-- ggplot(jewl_data, aes(x = Age, y = SpendingScore)) + geom_point(aes(color = h_cluster)),  -->
<!-- nrow  = 1, ncol = 2, common.legend = TRUE, legend = "right"),  -->
<!-- top = text_grob("Plotting Hierarchical Clusters")) -->


<!-- ``` -->


<!-- ```{r} -->
<!-- annotate_figure( -->
<!-- ggarrange( -->
<!-- ggplot(jewl_data, aes(x = Income, y = Savings)) + geom_point(aes(color = h_cluster)),  -->
<!-- ggplot(jewl_data, aes(x = Age, y = Savings)) + geom_point(aes(color = h_cluster)), -->
<!-- ggplot(jewl_data, aes(x = Savings, y = SpendingScore)) + geom_point(aes(color = h_cluster)),  -->
<!-- ggplot(jewl_data, aes(x = Age, y = Income)) + geom_point(aes(color = h_cluster)), -->
<!-- nrow  = 2, ncol = 2, common.legend = TRUE, legend = "right"),  -->
<!-- top = text_grob("Plotting Hierarchical Clusters")) -->
<!-- ``` -->




<!-- ```{r} -->

<!-- library(ggpubr) -->
<!-- ggscatter(jewl_data, x = "Income", y = "SpendingScore", -->
<!--           color = "h_cluster", palette = "npg", -->
<!--           shape = "h_cluster", ellipse = TRUE,  -->
<!--           mean.point = TRUE, star.plot = TRUE, -->
<!--           ggtheme = theme_minimal()) -->
<!-- ``` -->



<!-- ```{r} -->

<!-- # Grouped Scatter plot with marginal density plots -->
<!-- # ggscatterhist( -->
<!-- #   jewl_data, x = "Income", y = "SpendingScore", -->
<!-- #   color = "cluster", size = 5, alpha = 0.6, -->
<!-- #   palette = get_palette(c("#00AFBB", "#FC4E07"), 5)),  -->
<!-- #   margin.params = list(fill = "cluster", color = "black", size = 0.2) -->
<!-- # )  -->
<!-- #   -->

<!-- ggscatterhist( -->
<!--   jewl_data, x = "Income", y = "SpendingScore", -->
<!--   color = "h_cluster", size = 3, alpha = 0.6, -->
<!--   margin.params = list(fill = "h_cluster", color = "black", size = 0.2) -->
<!--   ) -->

<!-- ``` -->



<!-- ```{r} -->

<!-- jewl_data$h_cluster = as.numeric(jewl_data$h_cluster) -->
<!-- h_cluster = as.numeric(jewl_data$h_cluster) -->
<!-- fviz_cluster(list(data = jewl_data, cluster = h_cluster)) -->
<!-- ``` -->















